---
title: "SAR11 Protein Comparison"
date: "2023-05-09"
output:
  html_document:
    theme: cosmo
    highlight: pygments
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
---
# Project Information
Talk about it, why? 


# Additional Information:
RPubs Link for Euryarchaeota: https://rpubs.com/sggarcia6/1042560


# Load your libraries
```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(kableExtra)
library(DT)
library(reshape2)
library(Biostrings)
library(tm)
```

# BLAST Comparison of 45 (HTCC1062) to others.

Having the comparison of 45 to 95 into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
bltabl1 <- read.csv("../output/45-95_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
 bltabl1_2.0 <- bltabl1[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

bltabl1_2.0_df <- data_frame(bltabl1_2.0)

colnames(bltabl1_2.0_df) <- c("Query", "Database", "Eval", "Score")
```

Having the comparison of 45 to GCA_012276695.1(also known as NP1) into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
bltabl2 <- read.csv("../output/45-NP1_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
 bltabl2_2.0 <- bltabl2[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

bltabl2_2.0_df <- data_frame(bltabl2_2.0)

colnames(bltabl2_2.0_df) <- c("Query", "Database", "Eval", "Score")
```

Having the comparison of 45 to GCA_000155895 into a CSV file.  (HTCC7211)
```{r, eval=TRUE}
# Read out the blast file into csv
bltabl3 <- read.csv("../output/bigblast-01/GCA_000155895.1_45_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
 bltabl3_2.0 <- bltabl3[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

bltabl3_2.0_df <- data_frame(bltabl3_2.0)

colnames(bltabl3_2.0_df) <- c("Query", "Database", "Eval", "Score")
```

Having the comparison of 45 to GCA_002101295 into a CSV file. (Alphaproteobacteria	RS40)
```{r, eval=TRUE}
# Read out the blast file into csv
bltabl4 <- read.csv("../output/bigblast-01/GCA_002101295.1_45_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
 bltabl4_2.0 <- bltabl4[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

bltabl4_2.0_df <- data_frame(bltabl4_2.0)

colnames(bltabl4_2.0_df) <- c("Query", "Database", "Eval", "Score")
```

Having the comparison of 45 to GCA_028230015.1 into a CSV file. (AM-375-I16)
```{r, eval=TRUE}
# Read out the blast file into csv
bltabl5 <- read.csv("../output/bigblast-01/GCA_028230015.1_45_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
 bltabl5_2.0 <- bltabl5[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

bltabl5_2.0_df <- data_frame(bltabl5_2.0)

colnames(bltabl5_2.0_df) <- c("Query", "Database", "Eval", "Score")
```

Having the comparison of 45 to GCA_028229715.1 into a CSV file. (AM-375-K20)
```{r, eval=TRUE}
# Read out the blast file into csv
bltabl6 <- read.csv("../output/bigblast-01/GCA_028229715.1_45_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
 bltabl6_2.0 <- bltabl6[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

bltabl6_2.0_df <- data_frame(bltabl6_2.0)

colnames(bltabl6_2.0_df) <- c("Query", "Database", "Eval", "Score")
```

# Reading the file from 45 protein to CSV file.
```{r, eval=TRUE}
spgo <- read.csv("../output/45protein.tab", sep = '\t', header = TRUE)
colnames(spgo) <- c("Database", "Protein", "Coding_Sequence")
```

# Tables for each comparison
45 to 95
```{r, eval=TRUE}
# Join the data frames
joined1_df <- left_join(bltabl1_2.0_df, spgo, by = c("Database" = "Database")) %>%
  mutate(Organism = "sar95")

# You may print the table with kable
#kable(head(joined1_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 

#Scrollable tables
datatable(head(joined1_df), options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = FALSE))  
```
45 to GCA_012276695.1
```{r, eval=TRUE}
# Join the data frames
joined2_df <- left_join(bltabl2_2.0_df, spgo, by = c("Database" = "Database")) %>%
  mutate(Organism = "NP1")

# You may print the table with kable
#kable(head(joined2_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 

#Scrollable tables
datatable(head(joined2_df), options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = FALSE)) 
```
45 to GCA_000155895.1
```{r, eval=TRUE}
# Join the data frames
joined3_df <- left_join(bltabl3_2.0_df, spgo, by = c("Database" = "Database")) %>%
  mutate(Organism = "HTCC711")

# You may print the table with kable
#kable(head(joined3_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 

#Scrollable tables
datatable(head(joined3_df), options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = FALSE))  
```
45 to GCA_002101295.1
```{r, eval=TRUE}
# Join the data frames
joined4_df <- left_join(bltabl4_2.0_df, spgo, by = c("Database" = "Database")) %>%
  mutate(Organism = "RS40")

#You may print the table with kable
#kable(head(joined4_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 

#Scrollable tables
datatable(head(joined4_df), options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = FALSE))   
```
45 to GCA_028230015.1
```{r, eval=TRUE}
# Join the data frames
joined5_df <- left_join(bltabl5_2.0_df, spgo, by = c("Database" = "Database")) %>%
  mutate(Organism = "AM-375-I16")

# You may print the table with kable
#kable(head(joined5_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 

#Scrollable tables
datatable(head(joined5_df), options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = FALSE)) 
  
```
45 to GCA_028229715.1
```{r, eval=TRUE}
# Join the data frames
joined6_df <- left_join(bltabl6_2.0_df, spgo, by = c("Database" = "Database")) %>%
  mutate(Organism = "AM-375-K20")

# You may print the table with kable
#kable(head(joined6_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 

#Scrollable tables
datatable(head(joined6_df), options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = FALSE)) 
```
# Joining all Datatables 
```{r, eval=TRUE}
# Join the data frames
joinedtotal_df <- bind_rows(joined1_df, joined2_df, joined3_df, joined4_df, joined5_df, joined6_df)

#You may print the table with kable
#kable(head(joinedtotal_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 

#Scrollable tables
datatable(head(joinedtotal_df), options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = FALSE)) 
```

# Two different Heatmaps
```{r, eval=TRUE}

# Summarize the data
summarized_df <- joinedtotal_df %>%
  group_by(Protein, Organism) %>%
  summarize(mean_value = mean(Score))

# Create a wide-format table
wide_df <- summarized_df %>%
  pivot_wider(names_from = Organism, values_from = mean_value)

# Exclude non-numeric columns
data_mat <- as.matrix(wide_df[, -1])

# Define custom color palette
my_palette <- colorRampPalette(c("orange", "purple", "white"))(n = 100)

# Generate the heatmap with custom colors and row names
heatmap(data_mat, Rowv = NA, Colv = NA, col = my_palette, scale = "column",
        labRow = wide_df$Protein, main = "Heatmap Comparing 6", xlab = "Organism")

# Add color legend
color.legend <- function(color.palette, values, title = NULL, position = "topright") {
  n <- length(color.palette)
  x <- 0:(n - 1) / (n - 1)
  y <- 0.5
  rect(x, y, x + 1 / (n - 1), y + 1, col = color.palette, border = NA)
  text((2 * x + 1) / 2, y + 1.5, labels = values, adj = c(0.5, 0), cex = 0.8)
  if (!is.null(title)) {
    mtext(title, side = 1, line = 1, cex = 0.8)
  }
}
color.legend(color.palette = my_palette, values = c("Low", "Medium", "High"), title = "Legend")

# Adjust plot margins
par(mar = c(6, 6, 3, 4))



```
HeatMap comparing 6
```{r, eval=TRUE}
# Summarize the data
summarized_df <- joinedtotal_df %>%
  group_by(Protein, Organism) %>%
  summarize(mean_value = mean(Score))

# Create a wide-format table
wide_df <- summarized_df %>%
  pivot_wider(names_from = Organism, values_from = mean_value)

# Exclude non-numeric columns
data_mat <- as.matrix(wide_df[, -1])

# Define custom color palette
my_palette <- colorRampPalette(c("orange", "purple", "white"))(n = 100)

# Generate the heatmap with custom colors and row names
heatmap(data_mat, Rowv = NA, Colv = NA, col = my_palette, scale = "column",
        labRow = wide_df$Protein)
```

HeatMap comparing 4 with the top ten?
```{r, eval=TRUE}
# Summarize the data
summarized_df <- joinedtotal_df %>%
  group_by(Protein, Organism) %>%
  summarize(mean_value = mean(Score))

# Sort the summarized data by mean_value in descending order
summarized_df <- summarized_df %>%
  arrange(desc(mean_value))

# Select the top ten proteins
top_ten_proteins <- summarized_df$Protein[1:10]

# Filter the wide-format table to include only the top ten proteins
filtered_df <- wide_df %>%
  filter(Protein %in% top_ten_proteins)

# Exclude non-numeric columns
data_mat <- as.matrix(filtered_df[, -1])

# Define custom color palette
my_palette <- colorRampPalette(c("orange", "purple", "white"))(n = 100)

# Generate the heatmap with custom colors and row names
heatmap(data_mat, Rowv = NA, Colv = NA, col = my_palette, scale = "column",
        labRow = filtered_df$Protein)

```

# Top ten protein hits for all joined datasets

Joined Dataset 1 
```{r, eval=TRUE}
# Read joined blast table with annotation table
joined1_df <- left_join(bltabl1_2.0_df, spgo, by = "Database")
#c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined1_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

```


Joined Dataset 2
```{r, eval=TRUE}
# Read joined blast table with annotation table
joined2_df <- left_join(bltabl2_2.0_df, spgo, by = "Database")
#c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined2_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

```


Joined Dataset 3
```{r, eval=TRUE}
# Read joined blast table with annotation table
joined3_df <- left_join(bltabl3_2.0_df, spgo, by = "Database")
#c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined3_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

```

Joined Dataset 4
```{r, eval=TRUE}
# Read joined blast table with annotation table
joined4_df <- left_join(bltabl4_2.0_df, spgo, by = "Database")
#c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined4_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

```

Joined Dataset 5
```{r, eval=TRUE}
# Read joined blast table with annotation table
joined5_df <- left_join(bltabl5_2.0_df, spgo, by = "Database")
#c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined5_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

```

Joined Dataset 6
```{r, eval=TRUE}
# Read joined blast table with annotation table
joined6_df <- left_join(bltabl6_2.0_df, spgo, by = "Database")
#c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined6_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

```