---
title: "SAR11pres05"
format: revealjs
editor: visual
---

```{r}
library(knitr)
library(tidyverse)
library(kableExtra)
library(DT)
library(reshape2)
library(Biostrings)
library(tm)
```

## **SAR11 Bacteria Comparative Genomics Project**

SAR11 is a group of small, carbon-oxidizing bacteria that reach a global estimated population size of 2.4×10^28^ cells-approximately 25% of all plankton. They play a key role in the ocean's environment.

The goal of this comparative genomics project is to find presense and absence of SAR11 genomes. The genomes of interest were obtained from NCBI database.

## Methods

-   Determine goal of the project
-   Decided which genomes to use and obtained them from the NCBI database
-   Coding
    -   Make a database
    -   Get the query sequence
    -   Run BLAST
    -   Joining blast table with annotation table

The code was based on week 01 assignment from Fish546 Spring 23.

#Obtaining initial data

##Focused on two main genomes that were cultures and full genomes/

```{bash}
# Make a database

##Comparing protein sequence 012276695protein.faa (for easier terms called 95) to 000012345.1protein.faa (for easier terms called 45). First creating a database for 95.  

/home/shared/ncbi-blast-2.11.0+/bin/makeblastdb \
-in ../data_raw/012276695protein.faa \
-dbtype prot \
-out ../blastdb/95-protein
```

```{bash}
/home/shared/ncbi-blast-2.11.0+/bin/blastp \
-query ../data_raw/000012345.1protein.faa \
-db ../blastdb/95-protein \
-out ../output/95-45_blastp.tab \
-evalue 1E-20 \
-num_threads 20 \
-max_target_seqs 1 \
-outfmt 6
```

## Preliminary Results

Histogram for the Protein Sequences

```{r setup, include = TRUE}
# Read FASTA file
fasta_file <- "../data_raw/012276695protein.faa"  
psequences <- readDNAStringSet(fasta_file)

# Calculate sequence lengths
psequence_lengths <- width(psequences)

# Create a data frame
psequence_lengths_df <- data.frame(Length = psequence_lengths)

# Plot histogram using ggplot2
ggplot(psequence_lengths_df, aes(x = Length)) +
  geom_histogram(binwidth = 1, color = "purple", fill = "red", alpha = 0.75) +
  labs(title = "Histogram of P-Sequence Lengths",
       x = "P-Sequence Length",
       y = "Frequency") +
  theme_minimal()
```

Visual for Joining blast table with annotation table

```{r setup, include = TRUE}
# Read joined blast table with annotation table
joined_df <- left_join(bltabl2.0_df, spgo, by = c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()
```

## Next Steps:

-   fasta to tab-delimited for the protein sequences to finish the joining
-   figure out how this process can be done for multiple protein sequences from other SAR11 sequences
    -   which sequences to be chosen?
-   tyding up the figures that come from other pieces of code (understanding what the code does)
