---
title: "Exploring Euryarchaeota BLAST"
date: "2023-05-17"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
---
# Euryarchaeota in the Ocean

I was interested in seeing how similar would Marine Group lll Euryarchaeota would be to their neighbors in Euryarchaeota, specifically to Haloarchaea. I chose two reference genomes for to use for the BLAST. The first was DS2 or also known as Haloferax Volcanii and the second was Haloferax Marinum. I chose H. volcanii because it is a model for haloarchaea and I chose H marinum because it was low salt-tolerant haloarchaea isolated from seawater and sediment. Below I am attaching the protein names i used with names/origin of the organism.

## Respective Protein Names 
1. GCA_028346035.1: Marine Group III euryarchaeote 
Archaea; Candidatus Thermoplasmatota; Thermoplasmata; Candidatus Thermoprofundales.
2. GCA_028345155.1: Marine Group III euryarchaeote 
Archaea; Candidatus Thermoplasmatota; Thermoplasmata; Candidatus Thermoprofundales.
3. GCA_028343655.1: Marine Group III euryarchaeote 
Archaea; Candidatus Thermoplasmatota; Thermoplasmata; Candidatus Thermoprofundales.
4. GCA_012964945.1: Marine Group III euryarchaeote 
Archaea; Candidatus Thermoplasmatota; Thermoplasmata; Candidatus Thermoprofundales.
5. GCA_012963685.1: Marine Group III euryarchaeote 
Archaea; Candidatus Thermoplasmatota; Thermoplasmata; Candidatus Thermoprofundales.
6. GCA_012959635.1: Marine Group III euryarchaeote 
Archaea; Candidatus Thermoplasmatota; Thermoplasmata; Candidatus Thermoprofundales.
7. GCA_009674625.1: Haloferax Marinum
8. GCA_000025685.1: Haloferax Volcanii



# Loading Libraries

Remember to load your libraries

```{r setup, include = FALSE}
library(knitr)
library(tidyverse)
library(kableExtra)
library(DT)
library(reshape2)
library(Biostrings)
library(tm)
```

# Making a Database for DS2 or H.volcanii (Reference Genome)
```{r, engine='bash', eval = TRUE}
/home/shared/ncbi-blast-2.11.0+/bin/makeblastdb \
-in ../eury_data_raw/ds2_protein.faa \
-dbtype prot \
-out ../eury_blastdb/ds2-protein
```
# Making a Database for  H.marinum (Reference Genome)
```{r, engine='bash', eval = TRUE}
/home/shared/ncbi-blast-2.11.0+/bin/makeblastdb \
-in ../eury_data_raw/GCA_009674625.1_protein.faa \
-dbtype prot \
-out ../eury_blastdb/hmar-protein
```

# Exploring Files
## Finding the files within the folder
```{bash, eval = TRUE}
find ../eury_data_raw/ -type f -name "*.faa" -print0 | xargs -0 -I {} bash -c 'input_file="{}"; output_file="$(dirname "$input_file")/$(basename "$(dirname "$input_file")")_$(basename "$input_file")"; cp "$input_file" "$output_file"'

```
## How many files?
```{r, engine='bash', , eval = TRUE}
find ../eury_data_raw/ -type f -name "G*.faa" | wc -l
```
# BIG BLAST 
## Blasting all the files for DS2
```{r, engine='bash', eval = FALSE}

mkdir ../output/eury_bigblast-02

find ../eury_data_raw/ -type f -name "G*.faa" \
| xargs basename -s _protein.faa | xargs -I{} \
/home/shared/ncbi-blast-2.11.0+/bin/blastp \
-query ../eury_data_raw/{}/{}_protein.faa \
-db ../eury_blastdb/ds2-protein \
-out ../output/eury_bigblast-02/{}_ds2_blastp.tab \
-evalue 1E-20 \
-num_threads 40 \
-max_target_seqs 1 \
-outfmt 6
```
## Blasting all the files for H marinum
```{r, engine='bash', eval = FALSE}

#mkdir ../output/eury_bigblast-02

find ../eury_data_raw/ -type f -name "G*.faa" \
| xargs basename -s _protein.faa | xargs -I{} \
/home/shared/ncbi-blast-2.11.0+/bin/blastp \
-query ../eury_data_raw/{}/{}_protein.faa \
-db ../eury_blastdb/hmar-protein \
-out ../output/eury_bigblast-02/{}_hmar_blastp.tab \
-evalue 1E-20 \
-num_threads 40 \
-max_target_seqs 1 \
-outfmt 6
```

# Comparisons 
## Blasted to DS2
Comparison of DS2 to GCA_028346035.1 into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl1 <- read.csv("../output/eury_bigblast-02/GCA_028346035.1_ds2_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
eurybltabl1_2.0 <- eurybltabl1[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl1_2.0_df <- data_frame(eurybltabl1_2.0)

colnames(eurybltabl1_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_028345155.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl2 <- read.csv("../output/eury_bigblast-02/GCA_028345155.1_ds2_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
eurybltabl2_2.0 <- eurybltabl2[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl2_2.0_df <- data_frame(eurybltabl2_2.0)

colnames(eurybltabl2_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_028343655.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl3 <- read.csv("../output/eury_bigblast-02/GCA_028343655.1_ds2_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
eurybltabl3_2.0 <- eurybltabl3[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl3_2.0_df <- data_frame(eurybltabl3_2.0)

colnames(eurybltabl3_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_012964945.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl4 <- read.csv("../output/eury_bigblast-02/GCA_012964945.1_ds2_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
eurybltabl4_2.0 <- eurybltabl4[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl4_2.0_df <- data_frame(eurybltabl4_2.0)

colnames(eurybltabl4_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_012963685.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl5 <- read.csv("../output/eury_bigblast-02/GCA_012963685.1_ds2_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
eurybltabl5_2.0 <- eurybltabl5[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl5_2.0_df <- data_frame(eurybltabl5_2.0)

colnames(eurybltabl5_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_012960615.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl6 <- read.csv("../output/eury_bigblast-02/GCA_012960615.1_ds2_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
eurybltabl6_2.0 <- eurybltabl6[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl6_2.0_df <- data_frame(eurybltabl6_2.0)

colnames(eurybltabl6_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_012959635.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl7 <- read.csv("../output/eury_bigblast-02/GCA_012959635.1_ds2_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
eurybltabl7_2.0 <- eurybltabl7[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl7_2.0_df <- data_frame(eurybltabl7_2.0)

colnames(eurybltabl7_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_009674625.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl8 <- read.csv("../output/eury_bigblast-02/GCA_009674625.1_ds2_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
eurybltabl8_2.0 <- eurybltabl8[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl8_2.0_df <- data_frame(eurybltabl8_2.0)

colnames(eurybltabl8_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of DS2 to GCA_000025685.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
eurybltabl9 <- read.csv("../output/eury_bigblast-02/GCA_000025685.1_ds2_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
eurybltabl9_2.0 <- eurybltabl9[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

eurybltabl9_2.0_df <- data_frame(eurybltabl9_2.0)

colnames(eurybltabl9_2.0_df) <- c("Query", "Database", "Eval", "Score")
```


## Blasted to H marinum
Comparison of H marinum to GCA_028346035.1 into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl1 <- read.csv("../output/eury_bigblast-02/GCA_028346035.1_hmar_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
hmarbltabl1_2.0 <- hmarbltabl1[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl1_2.0_df <- data_frame(hmarbltabl1_2.0)

colnames(hmarbltabl1_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_028345155.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl2 <- read.csv("../output/eury_bigblast-02/GCA_028345155.1_hmar_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
hmarbltabl2_2.0 <- hmarbltabl2[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl2_2.0_df <- data_frame(hmarbltabl2_2.0)

colnames(hmarbltabl2_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_028343655.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl3 <- read.csv("../output/eury_bigblast-02/GCA_028343655.1_hmar_blastp.tab", sep = '\t', header = FALSE)


#Extract columns desired
hmarbltabl3_2.0 <- hmarbltabl3[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl3_2.0_df <- data_frame(hmarbltabl3_2.0)

colnames(hmarbltabl3_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_012964945.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl4 <- read.csv("../output/eury_bigblast-02/GCA_012964945.1_hmar_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
hmarbltabl4_2.0 <- hmarbltabl4[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl4_2.0_df <- data_frame(hmarbltabl4_2.0)

colnames(hmarbltabl4_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_012963685.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl5 <- read.csv("../output/eury_bigblast-02/GCA_012963685.1_hmar_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
hmarbltabl5_2.0 <- hmarbltabl5[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl5_2.0_df <- data_frame(hmarbltabl5_2.0)

colnames(hmarbltabl5_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_012960615.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl6 <- read.csv("../output/eury_bigblast-02/GCA_012960615.1_hmar_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
hmarbltabl6_2.0 <- hmarbltabl6[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl6_2.0_df <- data_frame(hmarbltabl6_2.0)

colnames(hmarbltabl6_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_012959635.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl7 <- read.csv("../output/eury_bigblast-02/GCA_012959635.1_hmar_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
hmarbltabl7_2.0 <- hmarbltabl7[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl7_2.0_df <- data_frame(hmarbltabl7_2.0)

colnames(hmarbltabl7_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_009674625.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl8 <- read.csv("../output/eury_bigblast-02/GCA_009674625.1_hmar_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
hmarbltabl8_2.0 <- hmarbltabl8[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl8_2.0_df <- data_frame(hmarbltabl8_2.0)

colnames(hmarbltabl8_2.0_df) <- c("Query", "Database", "Eval", "Score")
```
Comparison of H marinum to GCA_000025685.1  into a CSV file.
```{r, eval=TRUE}
# Read out the blast file into csv
hmarbltabl9 <- read.csv("../output/eury_bigblast-02/GCA_000025685.1_hmar_blastp.tab", sep = '\t', header = FALSE)

#Extract columns desired
hmarbltabl9_2.0 <- hmarbltabl9[, c("V1", "V2", "V11", "V12")]

#Make into dataframe

hmarbltabl9_2.0_df <- data_frame(hmarbltabl9_2.0)

colnames(hmarbltabl9_2.0_df) <- c("Query", "Database", "Eval", "Score")
```



# Setting up the Reference Genome Information needed

## DS2
DS2 fasta to tab-delimited
```{r, engine='bash'}
perl -e '$count=0; $len=0; while(<>) {s/\r?\n//; s/\t/ /g; if (s/^>//) { if ($. != 1) {print "\n"} s/ |$/\t/; $count++; $_ .= "\t";} else {s/ //g; $len += length($_)} print $_;} print "\n"; warn "\nConverted $count FASTA records in $. lines to tabular format\nTotal sequence length: $len\n\n";' ../eury_data_raw/ds2_protein.faa > ../output/ds2protein.tab
```

Reading the file for ds2 protein to CSV file
```{r, eval=TRUE}
ds2csv <- read.csv("../output/ds2protein.tab", sep = '\t', header = TRUE)
colnames(ds2csv) <- c("Database", "Protein", "Coding_Sequence")
```

## H marinum
H. marinum fasta to tab-delimited
```{r, engine='bash'}
perl -e '$count=0; $len=0; while(<>) {s/\r?\n//; s/\t/ /g; if (s/^>//) { if ($. != 1) {print "\n"} s/ |$/\t/; $count++; $_ .= "\t";} else {s/ //g; $len += length($_)} print $_;} print "\n"; warn "\nConverted $count FASTA records in $. lines to tabular format\nTotal sequence length: $len\n\n";' ../eury_data_raw/GCA_009674625.1_protein.faa > ../output/hmarprotein.tab
```

Reading the file for H. marinum protein to CSV file
```{r, eval=TRUE}
hmarcsv <- read.csv("../output/hmarprotein.tab", sep = '\t', header = TRUE)
colnames(hmarcsv) <- c("Database", "Protein", "Coding_Sequence")
```

# Tables for each comparison to DS2

DS2 to GCA_028346035.1
```{r, eval=TRUE}
# Join the data frames
euryjoined1_df <- left_join(eurybltabl1_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "035")

# Print the table with kable
kable(head(euryjoined1_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_028345155.1
```{r, eval=TRUE}
# Join the data frames
euryjoined2_df <- left_join(eurybltabl2_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "155")

# Print the table with kable
kable(head(euryjoined2_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_028343655.1
```{r, eval=TRUE}
# Join the data frames
euryjoined3_df <- left_join(eurybltabl3_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "655")

# Print the table with kable
kable(head(euryjoined3_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_012964945.1
```{r, eval=TRUE}
# Join the data frames
euryjoined4_df <- left_join(eurybltabl4_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "945")

# Print the table with kable
kable(head(euryjoined4_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_012963685.1
```{r, eval=TRUE}
# Join the data frames
euryjoined5_df <- left_join(eurybltabl5_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "685")

# Print the table with kable
kable(head(euryjoined5_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_012960615.1
```{r, eval=TRUE}
# Join the data frames
euryjoined6_df <- left_join(eurybltabl6_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "615")

# Print the table with kable
kable(head(euryjoined6_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_012959635.1
```{r, eval=TRUE}
# Join the data frames
euryjoined7_df <- left_join(eurybltabl7_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "635")

# Print the table with kable
kable(head(euryjoined7_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_009674625.1
```{r, eval=TRUE}
# Join the data frames
euryjoined8_df <- left_join(eurybltabl8_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "625")

# Print the table with kable
kable(head(euryjoined8_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
DS2 to GCA_000025685.1
```{r, eval=TRUE}
# Join the data frames
euryjoined9_df <- left_join(eurybltabl9_2.0_df, ds2csv, by = c("Database" = "Database")) %>%
  mutate(Organism = "685")

# Print the table with kable
kable(head(euryjoined9_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```

# Tables for each comparison to H marinum

H marinum to GCA_028346035.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined1_df <- left_join(hmarbltabl1_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "035")

# Print the table with kable
kable(head(hmarjoined1_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_028345155.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined2_df <- left_join(hmarbltabl2_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "155")

# Print the table with kable
kable(head(hmarjoined2_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_028343655.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined3_df <- left_join(hmarbltabl3_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "655")

# Print the table with kable
kable(head(hmarjoined3_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_012964945.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined4_df <- left_join(hmarbltabl4_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "945")

# Print the table with kable
kable(head(hmarjoined4_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_012963685.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined5_df <- left_join(hmarbltabl5_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "685")

# Print the table with kable
kable(head(hmarjoined5_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_012960615.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined6_df <- left_join(hmarbltabl6_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "615")

# Print the table with kable
kable(head(hmarjoined6_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_012959635.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined7_df <- left_join(hmarbltabl7_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "635")

# Print the table with kable
kable(head(hmarjoined7_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_009674625.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined8_df <- left_join(hmarbltabl8_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "625")

# Print the table with kable
kable(head(hmarjoined8_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```
H marinum to GCA_000025685.1
```{r, eval=TRUE}
# Join the data frames
hmarjoined9_df <- left_join(hmarbltabl9_2.0_df, hmarcsv, by = c("Database" = "Database")) %>%
  mutate(Organism = "685")

# Print the table with kable
kable(head(hmarjoined9_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
  
```

# Joining all Datatables for DS2
```{r, eval=TRUE}
# Join the data frames
ds2joinedtotal_df <- bind_rows(euryjoined1_df, euryjoined2_df, euryjoined3_df, euryjoined4_df, euryjoined5_df, euryjoined6_df, euryjoined7_df, euryjoined8_df, euryjoined9_df)

# Print the table with kable
kable(head(ds2joinedtotal_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)")
```

# Joining all Datatables for H marinum 
```{r, eval=TRUE}
# Join the data frames
hmar_joinedtotal_df <- bind_rows(hmarjoined1_df, hmarjoined2_df, hmarjoined3_df, hmarjoined4_df, hmarjoined5_df, hmarjoined6_df, hmarjoined7_df, hmarjoined8_df, hmarjoined9_df)

# Print the table with kable
kable(head(hmar_joinedtotal_df, n = 10), caption = "Joined data frame with kable formatting (first 10 rows)") 
```

# Visualization

## HeatMap comparing 9 for DS2

```{r, eval=TRUE}
# Summarize the data
ds2summarized_df <- ds2joinedtotal_df %>%
  group_by(Protein, Organism) %>%
  summarize(mean_value = mean(Score))

# Create a wide-format table
ds2wide_df <- ds2summarized_df %>%
  pivot_wider(names_from = Organism, values_from = mean_value)

# Exclude non-numeric columns
ds2data_mat <- as.matrix(ds2wide_df[, -1])

# Define custom color palette
ds2my_palette <- colorRampPalette(c("purple", "blue", "magenta"))(n = 100)

# Generate the heatmap with custom colors and row names
heatmap(ds2data_mat, Rowv = NA, Colv = NA, col = ds2my_palette, scale = "column",
        labRow = ds2wide_df$Protein)
```





## HeatMap comparing 9 for H mar

```{r, eval=TRUE}
# Summarize the data
hmar_summarized_df <- hmar_joinedtotal_df %>%
  group_by(Protein, Organism) %>%
  summarize(mean_value = mean(Score))

# Create a wide-format table
wide_df <- hmar_summarized_df %>%
  pivot_wider(names_from = Organism, values_from = mean_value)

# Exclude non-numeric columns
data_mat <- as.matrix(wide_df[, -1])

# Define custom color palette
my_palette <- colorRampPalette(c("orange", "purple", "pink"))(n = 100)

# Generate the heatmap with custom colors and row names
heatmap(data_mat, Rowv = NA, Colv = NA, col = my_palette, scale = "column",
        labRow = wide_df$Protein)
```

## Joined Dataset 1 Protein Hits Figure 
Joined Dataset 1 top Protein Hits
```{r, eval=TRUE}
# Read joined blast table with annotation table
joined1_df <- left_join(eurybltabl1_2.0_df, ds2csv, by = "Database")
#c("Database" = "Database"))

# Select the column of interest
column_name <- "Protein"  # Replace with the name of the column of interest
column_data <- joined1_df[[column_name]]

# Count the occurrences of the strings in the column
string_counts <- table(column_data)

# Convert to a data frame, sort by count, and select the top 10
string_counts_df <- as.data.frame(string_counts)
colnames(string_counts_df) <- c("String", "Count")
string_counts_df <- string_counts_df[order(string_counts_df$Count, decreasing = TRUE), ]
top_10_strings <- head(string_counts_df, n = 10)

# Plot the top 10 most common strings using ggplot2
ggplot(top_10_strings, aes(x = reorder(String, -Count), y = Count, fill = String)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Top 10 Protein hits",
       x = column_name,
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()

```